Supporting Information

### 2016-09-14 portfolio/text/ms.Rmd
Start drafting results

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
```

@@

Our models focused on the predictor species diversity,
while controlling for effort.
We calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of
Simpson's diversity index $\lambda$ [@simpson1949]
weighted by revenue $R$ ('species diversity'):

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ and
expected $\log(R_{i,t})$ is modelled as:

@@

>>>>

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

@@

We combined similar permits into permit groupings 
and identified sets of 
permit groupings held by an individual 
in a specific year as 'strategies' (Supporting Information). 
For example, we combined permits that were the same except for vessel size,
and we combined a number of permits that were targeting a single species 
and differed only in region fished.
Therefore, someone fishing herring roe in Kodiak
and someone fishing herring roe in Alaska Peninsula
were both considered to have a 'herring roe' permit group.
For salmon permits, 
we grouped permits using the same gear type across regions
for groups that shared similar species composition. 
After this process of combining permits,
we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}),
which we aggregated
within each person-year combination 
to form permit strategies.
For example, if someone fished halibut
and sablefish in the same year,
their strategy for that year would be `halibut-sablefish'.
Specific details on our strategy definitions are available in 
the Supporting Information.

Our models focused on species diversity
as a possible predictor of revenue and revenue variability
while controlling for effort.
We calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of
Simpson's diversity index $\lambda$ [@simpson1949]
weighted by revenue $R$ ('species diversity'):

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ 
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:

@@

\noindent
Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization in Alaska

- move towards specialization, salmon fig 1D (overall declining trend disappears if you exclude salmon)
- halibut most often combined with other permits (exclude permit holders that hold any halibut and about ninety percent of the remaining permit holders hold only one permit) fig 1c
- Salmon the most often on its own (exclude salmon and the proportioned single permit decreases to about 70%) fig 1c
- majority are single permit holders (70-90%) fig 1C, but the single permit holders make up only about half of the overall revenue (and surpassed half for the first time in 2013 and stayed that way) fig 1B

\noindent
Something about overall model fit

- Parameters from hierarchical model (main: Fig. S\ref{fig:fe}, varying: S\ref{fig:re})
- Residuals clearly show pattern that we see in the parameter estimates (Fig. S\ref{fig:resid})
- Downside residuals look very similar and justify using the full data set
(upside and downside) (Fig. S\ref{fig:resid-downside})
- Mention the year-permit varying intercepts and what they mean 
(Fig. S\ref{fig:str-yrs}) --- good and bad years where 
everyone in that strategy tended to do better or worse ---
this accounts for that and therefore compares everything to 
the expected revenue for that year

\noindent
Fig. \ref{fig:spokes} - add permits

- nearly universal reduction in variability (Fig. S\ref{fig:spoke-all})
- e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] 
lower expected standard deviation
for halibut plus another permit vs. single halibut permits);
- usually associated with higher median revenue 
- but can be very expensive (provide some examples from website)

\noindent
Fig. \ref{fig:across} - across strategies

- clear exponential decline in standardized variability with strategies that are inherently more species diverse 
- not a clear effect of days fished at this level
- fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);

\noindent
Fig. \ref{fig:within} - within strategies

- fish more days, make more money (Fig. S\ref{fig:fe})
- after accounting for days fished, most commonly, 
specializing is associated with higher revenue, 
but sometimes the reverse (Fig. \ref{fig:within})
- if we don't account for changes in effort, then higher species diversity 
means more money fairly consistently (TODO make figure)
- changing species diversity (number of species fished or evenness of species
fished) tends to be associated with greater variability / uncertainty
year-to-year (Figs. \ref{fig:within}, S\ref{fig:fe})
- increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
(Fig. \ref{fig:within})

### 2016-09-14 portfolio/text/supp.Rnw
Keep developing methods
Start drafting results

@@

We then proceeded to combine similar permits in the following steps.
We combined a number of permits that were the same except for vessel size.
We combined permits for
otter troll,
longline finfish,
pot gear.
king crab,
and
shrimp.
We grouped a number of permits that were targeting a single species
and differed only in region fished.
For example, someone fishing herring roe in the Kodiak (G34K)
and someone fishing herring roe in Alaska Peninsula (G34M)
were both considered to have a herring roe permit strategy.
We combined permits across each of
dungeness crab,
herring roe purse seining,
herring roe gillnet,
herring hand picked,
herring pound fishing,
sea cucumber,
sea urchin,
tanner crabs

@@

After this process of combining permits,
we were left with \Sexpr{n_grouped_permits}
unique fishing strategies (Table~\ref{tab:strategies}),
which we aggregated
within each person-year combination.
For example, if someone fished halibut
and sablefish in the same year
their strategy for that year would be ``halibut-sablefish''.
We removed people who switched strategies from one year to the next
since we were working with changes in revenue from year to year
and so needed consistent strategies across any pair of years
for a permit holder.
We removed strategies with less than 100 unique permit holders
across all years to focus on the major strategies
leaving us with a total of \Sexpr{nstr} strategies.
Finally, we removed all person-year combinations without
revenue recorded in the following year
in order to calculate percent change in revenue
from year to year.
This reduced the rows (person-year combinations)
in the final dataset from \Sexpr{ndat$annual} to
\Sexpr{ndat$annual_diff}.

@@

\section{Hierarchical model}

We checked our models with a two-part modelling approach
using the R package lme4.
To do this, we fit the initial model of the mean component $\log(R_{i,t})$
and then fit the second model to the log absolute value of the
residuals, $\log | \widehat{\log(R_{i,t})} - \log(R_{i,t}) |$.
Most importantly, (1) this model does not carry uncertainty
from the first mean-component model to the variance-component model, but
(2) it also leaves the residuals in the mean-component model strongly
heteroskedastic and therefore improperly weights individual data points, and
(3) does not allow for proper modelling of the strategy-level predictors,
$\eta_1$ and $\eta_2$. Nonetheless, it provides a check that our results
from the Stan model are approximately correct and allowed us
to quickly iterate and build intuition about the models.
A version of the Stan model without the variance-component
matched the estimates of the same model in lme4 almost exactly.

>>>>

\newlabel{eq:mean}{3}
\newlabel{eq:cv}{7}

@@

We then proceeded to combine similar permits in the following steps.
We combined a number of permits that were the same except for vessel size.
We combined permits for
otter troll,
longline finfish,
pot gear.
king crab,
and
shrimp.
We grouped a number of permits that were targeting a single species
and differed only in region fished.
For example, someone fishing herring roe in Kodiak (G34K)
and someone fishing herring roe in Alaska Peninsula (G34M)
were both considered to have a herring roe permit strategy.
We combined permits across each of
dungeness crab,
herring roe purse seining,
herring roe gillnet,
herring hand picked,
herring pound fishing,
sea cucumber,
sea urchin,
tanner crabs

@@

After this process of combining permits,
we were left with \Sexpr{n_grouped_permits}
unique fishing strategies (Table~\ref{tab:strategies}),
which we aggregated
within each person-year combination.
For example, if someone fished halibut
and sablefish in the same year
their strategy for that year would be `halibut-sablefish'.
We removed people who switched strategies from one year to the next
since we were working with changes in revenue from year to year
and so needed consistent strategies across any pair of years
for a permit holder.
We removed strategies with less than 100 unique permit holders
across all years to focus on the major strategies
leaving us with a total of \Sexpr{nstr} strategies.
Finally, we removed all person-year combinations without
revenue recorded in the following year
in order to calculate percent change in revenue
from year to year.
This reduced the rows (person-year combinations)
in the final dataset from \Sexpr{ndat$annual} to
\Sexpr{ndat$annual_diff}.

@@

\section{Model checking with two-stage approach}

We checked our models with a two-stage modelling approach
using the R package lme4 \citep{bates2015}.
To do this, we fit the initial model of the mean component $\log(R_{i,t})$
and then fit the second model to the log absolute value of the
residuals, $\log | \widehat{\log(R_{i,t})} - \log(R_{i,t}) |$.
Most importantly, (1) this model does not carry uncertainty
from the first mean-component model to the variance-component model, but
(2) it also leaves the residuals in the mean-component model strongly
heteroskedastic and therefore improperly weights individual data points, and
(3) does not allow for proper modelling of the strategy-level predictors,
$\eta_1$ and $\eta_2$. Nonetheless, it provides a check that our results
from the Stan model are approximately correct and allowed us
to quickly iterate and build intuition about the models.
A version of the Stan model without the variance-component
matched the estimates of the same model in lme4 almost exactly.
